\toolnameend: A Tool for Visualizing Patterns of Spreadsheet Function Combinations

### 2016-04-26 doc/paper.tex
Prep for SVG/PDF images.

\usepackage{textcomp}

@@

	\section{Introduction} Spreadsheets surround us. They serve key roles in
	industry, where as much as 95\% of U.S. financial firms use them daily
	\cite{panko2008sarbanes}, and academia, where teachers can use them to track
	research and students' grades, not to mention many other applications
	\cite{ko2011state}. It's this flexibility that grants them their allure: while
	a novice can still work without much experience, spreadsheets offer hundreds of
	specialized functions to fulfill a wide range of advanced needs, from complex
	statistics to text manipulation \cite{nardi1990spreadsheet}. As such, it should
	be no surprise that tens of millions of workers rely on these programs every
	day \cite{scaffidi2005estimating}.

	However, this ubiquity is not without danger: as more people adopt these tools
	and make larger and more complex sheets, the cost of failure grows too. For examples
	of this, the European Spreadsheet Risks Interest Group (EuSpRIG) documents
	many horror stories of spreadsheets gone 
	wrong\footnote{http://www.eusprig.org/horror-stories.htm}. One such story tells
	of a 2013 economics paper which drew connections between debt and
	national growth. Its findings shaped many political initiatives throughout the
	United States and Europe. However, other researchers soon discovered that a
	critical selection error in the research spreadsheets hampered the results and
	that, by correcting this error, the findings reversed completely! And this
	is only one of several accounts; even if most spreadsheet errors are relatively
	benign, poor spreadsheet practice can risk millions, if not billions, of
	dollars \cite{powell2009impact}.

@@

	This paper contributes such a tool, \toolname, that attempts to fulfill this by
	building on the aforementioned spreadsheet corpora. Working with spreadsheets
	in Excel, \toolname focuses primarily on the functions, like SUM and IF and
	hundreds more, which are the building blocks of formula and programs which
	determine values across a spreadsheet. The tool creates an interactive,
	exploratory visualization which captures how people combine these functions in
	a selected collection of spreadsheets, quantifying the broad patterns of
	function nesting while also being specific enough to find anomalous formulae
	and link them back to the specific cell in spreadsheets where they originate.
	\par

@@

	To round out that discussion, we also conducted a user study to judge how
	usable and intuitive the tool actually is. This grounds the discussion of its
	potential uses while also underscoring some of its inherent limitations and
	setting the way for future extensions and improvements upon the concept.

>>>>

\usepackage{textcomp} 
\usepackage{svg}

@@

	\section{Introduction} Spreadsheets surround us. They serve key roles in
	industry, where as much as 95\% of U.S. financial firms use them daily
	\cite{panko2008sarbanes}, and academia, where teachers can use them to track
	research and students' grades, not to mention many other applications
	\cite{ko2011state}. It's this flexibility that grants them their allure: while
	a novice can still work without much experience, spreadsheets offer hundreds of
	specialized functions to fulfill a wide range of advanced needs, from complex
	statistics to text manipulation \cite{nardi1990spreadsheet}. As such, it should
	be no surprise that tens of millions of workers rely on these programs daily \cite{scaffidi2005estimating}.

	However, this ubiquity is not without danger; as more people adopt these tools
	and make larger and more complex sheets, the cost of failure grows too. To illustrate this, the European Spreadsheet Risks Interest Group (EuSpRIG) documents
	many horror stories of spreadsheets gone 
	wrong\footnote{http://www.eusprig.org/horror-stories.htm}. One story tells
	of a 2013 economics paper which drew connections between debt and
	national growth. Its findings shaped many political initiatives throughout the
	United States and Europe. However, other researchers soon discovered that a
	critical selection error in the research spreadsheets hampered the results and
	that, by correcting this error, the findings reversed completely! And this
	is only one of several accounts; even if most spreadsheet errors are relatively
	benign, poor spreadsheet practice can risk millions, if not billions, of
	dollars \cite{powell2009impact}.

@@

	This paper contributes such a tool, \toolname, that attempts to fulfill this
	need for better exploration by
	building on the aforementioned spreadsheet corpora. Working with spreadsheets
	in Excel, \toolname focuses primarily on the functions, like SUM and IF and
	hundreds more, which are the building blocks of formula and programs that
	determine values across a spreadsheet. The tool creates an interactive,
	exploratory visualization which captures how people combine these functions in
	a selected collection of spreadsheets, quantifying the broad patterns of
	function nesting while also being specific enough to find anomalous formulae
	and link them back to the specific cell in spreadsheets where they originate.
	\par

@@

	To better understand how people might use \toolnameend, we also conducted a user study to judge how
	usable and intuitive the tool actually is. This grounds the discussion of its
	potential uses while also underscoring some of its inherent limitations and
	setting the way for future extensions and improvements upon the concept.
